Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-28 sections/01-introduction.tex
Correct some typos and add covid prob assessment

Forecasting methods predict the future based on historical data, patterns, and exogenous factors. 
In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable. 
Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules \citep[see the review in][]{hyndman2006another}. 
They are not capable of assessing the trending ability discussed above.
In a prediction competition on armed conflicts, trending recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with a trending component for evaluation. 
However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
This paper focuses on pure trending evaluation that accompanies traditional scores.

Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \citep{browning1989nowcasting,schmid2019nowcasting}. 
%In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported \citep{gunther2021nowcasting}. 
Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay. 
Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \citep{castle2017forecasting}. 
Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \citep{gunther2021nowcasting} or probabilistic scoring rules \citep{Wolffram2023}. 
The aspect of trending ability is not considered in the literature to the best of our knowledge. 
However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.

Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors. 
A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \citep{watson2010method}. 
Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}. 
Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \citep{Saugel2015,saugel2018error,hiraishi2021concordance}.

@@

The remainder of this paper is structured as follows. 
Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries. 
In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
We conclude the paper in Section~\ref{sec:conclusion}.

>>>>

Forecasting methods predict the future based on historical data, patterns, and exogenous factors. 
In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable. 
Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules~\citep[see the review in][]{hyndman2006another}.
They are not capable of assessing the trending ability discussed above.
In a prediction competition on armed conflicts, trending recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with a trending component for evaluation. 
However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
This paper focuses on pure trending evaluation that accompanies traditional scores.

Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
%In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{gunther2021nowcasting}.
Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay. 
Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period~\citep{castle2017forecasting}.
Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse}~\citep{gunther2021nowcasting} or probabilistic scoring rules~\citep{Wolffram2023}.
The aspect of trending ability is not considered in the literature to the best of our knowledge. 
However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.

Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors. 
A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{watson2010method}.
Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}. 
Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \citep{Saugel2015,saugel2018error,hiraishi2021concordance}.

@@

The remainder of this paper is structured as follows.
Section~\ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts.
We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries.
In Section~\ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting.
Hence, we provide blueprints for applying trending evaluation in practice and make remarks on interpreting the measures.
We conclude the paper in Section~\ref{sec:conclusion}.

### 2024-03-28 sections/03-trending.tex
Correct some typos and add covid prob assessment

The definition of the \textit{predicted} change depends on the context;
Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$. 
We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
The predicted change is computed by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.

@@

@@

In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
However, systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.

@@

Probabilistic predictions can take the form of a \ac{cdf}, \ac{pdf}, or quantiles.
The \ac{cdf} is the most general and can be used to derive the others, given they exist.
Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
In the appendix, we extend the analysis for predictions given as quantiles or unknown true values.

>>>>

The definition of the \textit{predicted} change depends on the context;
Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$. 
We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
The predicted change is computed by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet.
If the true value $y_{t-l}$ is known at time $t$, the predicted change is computed by the difference between the nowcast and the true value, as $y_{t-l}$ is also known by the nowcaster and incorporated into the nowcast.
Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.

@@

In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
Systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
Note that in the case of measurement data, one missing value in the time series leads to two undefined differences in the change series; that is if $x_t$ is missing, $\diffxt$ and $\diffxt[t+l]$ are undefined; if an observation $y_t$ is missing, $\diffyt$ and $\diffyt[t+l]$ are undefined.
The data pair is excluded even if the corresponding nowcast or forecast is available.

@@

@@

Probabilistic predictions can take the form of a \ac{cdf}, \ac{pdf}, or quantiles.
The \ac{cdf} is the most general and can be used to derive the others, given they exist.
Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
In Appendix~\ref{subsec:app-probabilistic-trending-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.

### 2024-03-28 sections/04-application.tex
Correct some typos and add covid prob assessment

The data contains eight nowcasts from scientific and public institutions, nowcast communities, and a newspaper, using different input variables, calendar data, and length of training data.
The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
For information on the model design, we refer to~\citet{Wolffram2023}.
Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
We denote them by ENS-MEAN and ENS-MED.
\citet{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study. 
In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
We use the data from February 8, 2024, for the true values.
The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\citep{Tolksdorf2022}.

@@

In addition to close inspection of the point evaluation measures, assessing the trending of the nowcasts is crucial.
To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
Asymmetries are especially of interest in assessing whether some models are better at recognizing a fall than a rise or vice versa.

@@

Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different horizons.
Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
Overall, the variability and general level of differences grow with the horizon.
The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
Similarly, the 10\%-quantile of differences increases.
The 10\%-quantile is used for the exclusion areas in the trending assessment.
The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.

@@

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days. The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values. The measures with the exclusion area are denoted with the subscript $q_{0.1}$. }
    \label{tab:app-covid-trending-ratios-lag-7}
\end{table}

@@

Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas. 
This indicates that if the model predicts a large change, the direction is indeed better than when a small change is predicted.

@@

In a second example, we consider forecasting the number of arrivals in a large emergency department per hour.
Good forecasts are crucial for planning staff and resources.
Several models are used for predicting hourly outcomes in a study by \citet{Rostami-Tabar2023}.
Every 12 hours, the models issue hourly forecasts for the next 48 hours.
Thus, management can take measures according to the expected number of arrivals, for example, by redeploying staff and reconfiguring units.

@@

We start by analyzing the marginal distributions for the changes of forecasts and true values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
Thereby, the computed difference aligns with Section~\ref{subset:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
The variability of differences decreases for the larger horizon for most models; only for the ETS model does it increase.
The 10\%-quantile of the differences is between zero and one for all models and horizons.
Thus, we exclude only differences smaller than one from the trending assessment.
The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
\todo{Schwer verstaendlich (Auch Abbildung~\ref{tab:app-eda-marginals})}

@@

Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
The conditional trending plots show similar courses for the two horizons, while the curves are shifted downwards for the horizon of seven days.
The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
The GBM-2 model outperforms the qreg-1 model for all $x$.
The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
Benchmark-1 lies between the other models for all $x$.

Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
Remember that low a \ac{bs} and a plot along the diagonal are signs for good trending.
The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
The qreg-1 model has both horizons' highest \ac{bs}.
The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
For the other models, the reliability diagrams show a fraction of increases that are too large for the corresponding predicted probability.

@@

The four-quadrant plots contain a considerable number of extreme points.
Whether these points are due to measurement errors or extreme values is not distinguishable.
Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile \citep[see][]{Critchley2010}.
We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.

>>>>

The data contains eight nowcasts from scientific and public institutions, nowcast communities, and a newspaper, using different input variables, calendar data, and length of training data.
The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
For information on the model design, we refer to~\citet{Wolffram2023}.
In addition to those nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
We denote them by ENS-MEAN and ENS-MED.
\citet{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study.
In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
We use the data from February 8, 2024, for the true values.
The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\citep{Tolksdorf2022}.

@@

In addition to close inspection of the point evaluation measures, assessing the trending of the nowcasts is crucial.
To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
Asymmetries are especially relevant for assessing whether some models are better at recognizing a fall than a rise or vice versa.

@@

Before stating the results, we provide background information on the marginal distributions of the actual values and nowcasts for the different horizons.
Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
Overall, the variability and general level of differences grow with the horizon.
The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
Similarly, the 10\%-quantile of differences increases.
The 10\%-quantile is used for the exclusion areas in the trending assessment.
The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days. The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values. The subscript $q_{0.1}$ denotes the measures with exclusion area. }
    \label{tab:app-covid-trending-ratios-lag-7}
\end{table}

@@

Figure~\ref{fig:app-covid-prob} shows the \acf{bs} and reliability diagrams for the same subset of models, the ILM, RKI, RIVM, and ENS-MED.
The probabilities of increase for the different models are computed using the nowcast quantiles.
For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-trending-evaluation}).
Remember that low a \ac{bs} and a plot along the diagonal are signs of good trending.
The \ac{bs} is the lowest for the RIVM model, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
The RKI model yields the highest \ac{bs} for all horizons.
Note that the \ac{bs} is bounded by 0.25 for random guessing; thus, all models perform better than random guessing.
The reliability diagrams show that the models are not well calibrated for the horizon of one day.
While for predicted probabilities below 0.5, the observed ratio of increases is smaller, it is higher than predicted for probabilities above 0.5.
Figure~\ref{fig:app-covid-prob-hist} shows the observed predicted increase probabilities for all horizons.
For the horizons of seven and 14 days, the nowcasters issue only a few moderate probabilities.

\begin{figure}
    \begin{subfigure}{0.48\textwidth}
        \centering
        \tiny
        \input{plots/covid_nowcast/70_brier_scores}
    \caption{BS for the different models and horizons.}\label{fig:app-covid-prob-brier}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_1}
    \caption{Reliability diagram for horizon one day.}\label{fig:app-covid-prob-rel-1}
    \end{subfigure}\hspace{0.01\textwidth}%
    \caption[Brier Scores and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.
    The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
    In the computation of BS and reliability diagram, missing values are excluded.
    The reliability diagram for the horizon seven and 14 days is in the appendix (see Figure~\ref{fig:app-covid-rel}).}
    \label{fig:app-covid-prob}
\end{figure}

@@

Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas.
This indicates that if the model predicts a large change, the direction is more accurate than when a small change is predicted.

The probabilistic trending assessment shows that the models are better than random guessing.
In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
Thus, the reliability diagrams for the horizons of 7 and 14 days do not contain information on moderate probabilities.
The \ac{bs} values, however, work well for those examples and provide a good measure for the trending ability of the models.

@@

In a second example, we consider forecasting the number of arrivals in a large emergency department per hour.
Good forecasts are crucial for planning staff and resources.
In a study by \citet{Rostami-Tabar2023}, several models predict hourly outcomes.
Every 12 hours, the models issue hourly forecasts for the next 48 hours.
Thus, management can take measures according to the expected number of arrivals, for example, by redeploying staff and reconfiguring units.

@@

We start by analyzing the marginal distributions for the changes of forecasts and actual values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
Thereby, the computed difference aligns with Section~\ref{subset:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
The variability of differences decreases for the larger horizon for most models; only for the ETS model does it increase.
The 10\%-quantile of the differences is between zero and one for all models and horizons.
Thus, we exclude differences smaller than one from the trending assessment.
The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.

@@

Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
The conditional trending plots show similar courses for the two horizons, though the curves are shifted downwards for the horizon of seven days.
The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
The GBM-2 model outperforms the qreg-1 model for all $x$.
The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
Benchmark-1 lies between the other models for all $x$.

Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
The qreg-1 model has both horizons' highest \ac{bs}.
The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
For the other models, the reliability diagrams show a fraction of increases that are too large for the corresponding predicted probability.

@@

The four-quadrant plots contain a considerable number of extreme points.
Whether these points are due to measurement errors or extreme values is not distinguishable.
Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile~\citep[see][]{Critchley2010}.
We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.

### 2024-03-28 sections/05-conclusion.tex
Correct some typos and add covid prob assessment

Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
The positive and negative trending ratio analyze the trending ratio given the predicted change is positive or negative, respectively.
Thus, they quantify the credibility of the respective predictions. 
The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
In the applications, the bootstrap confidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or assess models with high trending differences.

>>>>

Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
The positive and negative trending ratios analyze the trending ratio given whether the predicted change is positive or negative, respectively.
Thus, they quantify the credibility of the respective predictions.
The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
In the applications, the bootstrap confidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high trending differences.
